---
title: "Lab 04 - Data Visualization"
output:
  html_document:
    html_preview: false
link-citations: yes
---



```{r setup, message=FALSE, warning=FALSE}
#install.packages(c("data.table","leaflet"))
library(data.table)
library(leaflet)
library(tidyverse)
```

# Learning Goals

- Read in and prepare the meteorological dataset
- Create several graphs with different `geoms()` in `ggplot2`
- Create a facet graph
- Conduct some customizations of the graphs
- Create a more detailed map using `leaflet()`


# Lab Description

We will again work with the 2023 meteorological data from last week.

**The objective of the lab is to examine the association between weekly average dew point temperature and wind speed in four regions of the US and by elevation.**

# Steps

### 1. Read in the data

First download and then read in with data.table:fread()

```{r checking-file}
fn <- "https://raw.githubusercontent.com/JSC370/JSC370-2024/main/data/met_all_2023.gz"
if (!file.exists("met_all_2023.gz"))
  download.file(fn, destfile = "met_all_2023.gz")
met <- data.table::fread("met_all_2023.gz")
```
```{r}
# Some EDA codes from last week
met$lat <- met$lat/1000
met$lon <- met$lon/1000
met$wind.sp <- met$wind.sp/10
met$temp <- met$temp/10
met$dew.point <- met$dew.point/10
met$atm.press <- met$atm.press/10

# Relative Humidity
met$rh <- 100*((112-0.1*met$temp+met$dew.point)/(112+0.9*met$temp))^8
```

### 2. Prepare the data

- Based on last week's lab, remove suspect/incorrect observations.
- Generate a date variable.
- Keep only the observations of the first week of the month.

In a new dataframe:
- Compute the mean by station (`"USAFID"`) of the variables `temp`, `rh`, `wind.sp`, `vis.dist`, `dew.point`, `lat`,`lon`, and `elev`.
- Create a categorical variable `elev_cat` for elevation (break point at 252 m)
- Create a categorical variable `region` for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
(Tip: you can use the `ifelse()` function in R)

```{r}
# write your code here


```


```{r, eval=FALSE}
# check number of stations in each region
table(met_avg$region, useNA = "always")
```

### 3. Use `geom_violin` to examine the wind speed and dew point temperature by region

You saw how to use `geom_boxplot` in class. Try using `geom_violin` instead (take a look at the help).
(hint: You will need to set the `x` aesthetic to 1)

- Use facets.
- Make sure to deal with `NA` category.
- Describe what you observe in the graph.

```{r}
# write your code here

```


### 4. Use `geom_jitter` with `stat_smooth` to examine the association between dew point temperature and wind speed by region

- Color points by region
- Make sure to deal with `NA` category
- Fit a linear regression line by region
- **Describe** what you observe in the graph

```{r}
# write your code here
  
```


### 5. Use `geom_bar` to create barplots of the weather stations by elevation category coloured by region

- Bars by elevation category using `position = "dodge"`
- Change colors from the default. Color by region using `scale_fill_brewer` see [this](http://rstudio-pubs-static.s3.amazonaws.com/5312_98fc1aba2d5740dd849a5ab797cc2c8d.html)
- Create nice labels on axes and add a title
- Try a different theme
- Describe what you observe in the graph
- Make sure to deal with `NA`

```{r}
# write your code here

```

### 6. Use `stat_summary` to examine mean dew point and wind speed by region with standard deviation error bars

- Make sure to remove `NA`
- Use fun.data="mean_sdl" in `stat_summary`
- Add another layer of `stats_summary` but change the geom to `"errorbar"` (see the help).
- Describe the graph and what you observe

```{r}
# write your code here

```

- Dew point temperature is...
- Wind speed is...

### 7. Make a map showing the spatial trend in relative humidity (`rh`) in the US

- Make sure to remove `NA`
- Use `leaflet()`
- Make a colour palette with custom colours
- Use `addMarkers` to include the top 10 places in relative h (hint: this will be useful `rank(-rh) <= 10`)
- Add a legend
- Describe trend in RH across the US

```{r}
# write your code here

```


### 8. Use a ggplot extension

- Pick and extension (except cowplot) from [here](https://exts.ggplot2.tidyverse.org/gallery/) and make a plot of your choice using the `met` data (or `met_avg`)
- Might want to try examples that come with the extension first (e.g. `ggtech`, `gganimate`, `ggforce`)

