---
title: "Lab 08 - Text Mining/NLP"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(eval = F, include  = T)
```

# Learning goals

- Use `unnest_tokens()` and `unnest_ngrams()` to extract tokens and ngrams from text
- Use dplyr and ggplot2 to analyze and visualize text data
- Try a theme model using `topicmodels`

# Lab description

For this lab we will be working with the medical record transcriptions from https://www.mtsamples.com/. And is loaded and "fairly" cleaned at https://github.com/JSC370/JSC370-2024/tree/main/data/medical_transcriptions.

This markdown document should be rendered using `github_document` document.



### Setup packages

You should load in `dplyr`, (or `data.table` if you want to work that way), `ggplot2` and `tidytext`. Install `wordcloud`, `tm`, and `topicmodels` if you don't alreadyh have them.



### Read in the Medical Transcriptions

Loading in reference transcription samples from https://www.mtsamples.com/

```{r eval=FALSE}
library(tidytext)
library(tidyverse)
library(wordcloud)
library(tm)
library(topicmodels)

mt_samples <- read_csv("https://raw.githubusercontent.com/JSC370/JSC370-2024/main/data/medical_transcriptions/mtsamples.csv")
mt_samples <- mt_samples %>%
  select(description, medical_specialty, transcription)

head(mt_samples)
```

---

## Question 1: What specialties do we have?

We can use `count()` from `dplyr` to figure out how many different medical specialties are in the data. Are these categories related? overlapping? evenly distributed? Make a bar plot.

```{r eval=FALSE}
mt_samples |>
  count(, sort = TRUE) |>
  ggplot()
```

---

## Question 2: Tokenize

- Tokenize the the words in the `transcription` column
- Count the number of times each token appears
- Visualize the top 20 most frequent words with a bar plot
- Create a word cloud

Explain what we see from this result. Does it makes sense? What insights (if any) do we get?

```{r eval=FALSE}
tokens <- mt_samples |>
  select(transcription) |>
  unnest_tokens() |>
  ...

tokens |>
  ggplot(aes()) +
  geom_bar()

wordcloud()
```

---

## Question 3: Stopwords

- Redo Question 2 but remove stopwords(use stopwords package)
- Bonus points if you remove numbers as well (use regex)

What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?

- Try customizing your stopwords list to include 3-4 additional words that do not appear informative

```{r eval=FALSE}
head(stopwords("english"))
length(stopwords("english"))


tokens <- mt_samples |>
  select(transcription) |>
  unnest_tokens() |>
  filter() |>
  ...

tokens |>
  ggplot(aes()) +
  geom_bar()

wordcloud()


```


---



# Question 4: ngrams

Repeat question 2, but this time tokenize into bi-grams. How does the result change if you look at tri-grams? Note we need to remove stopwords a little differently. You don't need to recreate the wordclouds.

```{r eval=FALSE}
sw_start <- paste0("^", paste(stopwords2, collapse=" |^"), "$")
sw_end <- paste0("", paste(stopwords2, collapse="$| "), "$")

tokens_bigram <- mt_samples |>
  select(transcription) |>
  unnest_tokens(ngram, transcription, token = "ngrams", n = 2) |>
  filter()|>
  filter()|>
  filter()|>
  ... 
```
---

# Question 5: Examining words

Using the results from the bigram, pick a word and count the words that appear before and after it, and create a plot of the top 20.

```{r eval=FALSE}
library(stringr)
# e.g. patient, blood, preoperative...
tokens_bigram |>
  filter(str_detect(#use regex for your word picking words before and after)) |>
    mutate(word = str_remove(ngram, ""),
         word = str_remove_all(word, " ")) |>
      ...
```

---


# Question 6: Words by Specialties

Which words are most used in each of the specialties. you can use `group_by()` and `top_n()` from `dplyr` to have the calculations be done within each specialty. Remember to remove stopwords. How about the most 5 used words?


```{r eval=FALSE}
mt_samples |>
   unnest_tokens(word, transcription) |>
   ...
   group_by(medical_specialty) |>
   ...
```


# Question 7: Topic Models

See if there are any themes in the data by using a topic model (LDA). 

- you first need to create a document term matrix
- then you can try the LDA function in `topicmodels`. Try different k values.
- create a facet plot of the results from the LDA (see code from lecture)


```{r eval=FALSE}

transcripts_dtm <- mt_samples |>
  select(transcription) |>
  unnest_tokens() |>
  filter() |>
  filter() |>
  DocumentTermMatrix()


transcripts_dtm <- as.matrix(transcripts_dtm)   

transcripts_lda <- LDA()

```



# Deliverables

1. Questions 1-7 answered, raw .Rmd file and pdf or html output uploaded to Quercus
