---
title: "Lab 03 - Exploratory Data Analysis"
output:
  tufte::tufte_html:
    css: style.css
link-citations: yes
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
#install.packages(c("data.table","leaflet"))
library(data.table)
library(leaflet)
library(tidyverse)
```

# Learning Goals

- Read in and get familiar with the meteorology dataset  
- Step through the EDA "checklist" presented in the class slides
- Practice making exploratory graphs

# Lab Description

We will work with the meteorological data presented in lecture, but updated to June 2023. Recall the dataset consists of weather station readings in the continental US. It is a large dataset so please use data.table throughout!

**The objective of the lab is to find the weather stations with the highest and lowest elevations and look at patterns in their time series of temperature and wind.speed**

# Steps

### 1. Read in the data

First download and then read in with data.table:fread()

```{r checking-file, include=FALSE}
fn <- "https://raw.githubusercontent.com/JSC370/JSC370-2024/main/data/met_all_2023.gz"
if (!file.exists("met_all_2023.gz"))
  download.file(fn, destfile = "met_all_2023.gz")
met <- data.table::fread("met_all_2023.gz")
```


```{r, echo=TRUE, message=FALSE, eval=FALSE}
download.file(
  "https://raw.githubusercontent.com/JSC370/JSC370-2024/main/data/met_all_2023.gz",
  destfile = "met_all_2023.gz",
  method   = "curl",
  timeout  = 60
  )

met <- data.table::fread("met_all_2023.gz")
```

### 2. Check the dimensions, headers, footers. How many columns, rows are there?

```{r}
dim(met)
head(met)
tail(met)
```


### 3. Take a look at the variables.

```{r}
str(met)
```


### 4. Take a closer look at the key variables. 

```{r}
table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
summary(met$lat)
summary(met$lon)
```

It looks like some of the variables are not scaled properly. We should take a deeper look at the data dictionary to confirm. 

After checking the data we should make the appropriate modifications. Specifically: divide lat and lon by 1000, divide wind.sp, temp, dew.point, atm.press by 10, and let's calculate relative humidity with the following: `100*((112-0.1*temp+dew.point)/(112+0.9*temp))^8`

```{r}
# <place your code here>
```

After scaling, check for implausible values of the key variables: temperature and humidity. Recall relative humidity should between 0% and 100%, and that these data are for June 2023.

Also check rates of missing values in the key variables.

```{r}
# <place your code here>
```

- _Summarize here_



### 5. Check the data against an external data source.

We should check the suspicious values (where is it located?) and validate that the range of elevations make sense. Google is your friend here.

Fix any problems that arise in your checks.

```{r}
# <place your code here>
```

- _Summarize anything that was removed_

### 6. Find the weather stations with the highest and lowest elevations

Using data.table you can order the data and print the head and tail.

After finding these sites make a `leaflet` map showing where these two sites are.

```{r}
library(leaflet)
# <place your code here>
```



### 7. Exploratory graphs

We should look at the distributions of all of the key variables (elevation, temp, wind speed) to make sure there are no remaining issues with the data.

```{r }
# <place your code here>
```

One thing we should consider for later analyses is to log transform wind speed and elevation as the are very skewed.


```{r}
# <place your code here>
```

- _Summarize_

Look at the time series of temperature and wind speed at this location. For this we will need to create a date-time variable for the x-axis.

```{r message=FALSE}
library(lubridate)
# <place your code here>
```

With the date-time variable we can plot the time series of temperature and wind speed. For the sake of computing time, do this for the weather station with the highest elevation.

```{r}
# <place your code here>
```

Summarize any trends that you see in these time series plots.

### 7. Look at basic relationships between variables

Remember to keep the initial question in mind. We want to pick out the weather station with maximum and minimum elevation and examine their wind speed and temperature.

Some ideas: 
1. calculate the correlation between temperature and wind speed; and 
2. make a scatterplot of temperature and wind speed and add the linear regression line (also look at `lm()` results)

```{r}
# <place your code here>
```
