---
title: "Lab 10 - Trees, Bagging, RF, Boosting XGBoost"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(eval = F, include  = T, echo = T)
```

# Learning goals

- Perform classification and regression with tree-based methods in R
- Recognize that tree-based methods are capable of capturing non-linearities by splitting multiple times on the same variables
- Compare the performance of classification trees, bagging, random forests, and boosting for predicting heart disease based on the ``heart`` data.

# Lab description

For this lab we will be working with the `heart` dataset that you can download from [here](https://github.com/JSC370/JSC370-2024/blob/main/data/heart.csv)


### Setup packages

You should install and load `rpart` (trees), `randomForest` (random forest), `gbm` (gradient boosting) and `xgboost` (extreme gradient boosting).


```{r}
install.packages(c("rpart", "rpart.plot", "randomForest", "gbm", "xgboost"))
```

### Load packages and data
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)
library(xgboost)
library(caret)

heart <- read_csv("https://raw.githubusercontent.com/JSC370/JSC370-2024/main/data/heart.csv")

  mutate(
    AHD = 1 * (AHD == "Yes"),
    ChestPain = factor(ChestPain),
    Thal = factor(Thal)
  )

head(heart)
```


---


## Question 1: Trees
- Split the `heart` data into training and testing (70-30%)

```{r}

train<-sample(1:nrow(heart), round(0.7*nrow(heart)))
heart_train<-heart[train,]
heart_test<-heart[-train,]
```

- Fit a classification tree using rpart, plot the full tree. We are trying to predict AHD. Set minsplit = 10, minbucket = 3, and do 10 cross validations.
```{r}
heart_tree <- rpart(
)
rpart.plot(heart_tree)
```

- Plot the complexity parameter table for an rpart fit and find the optimal cp
```{r}
plotcp(heart_tree)
printcp(heart_tree)

optimalcp = # extract from cptable
```

- Prune the tree

```{r}
heart_tree_prune<-prune()
rpart.plot()
```

- Compute the test misclassification error

```{r}
heart_pred<-predict(heart_tree_prune, heart_test)
...
```

- Fit the tree with the optimal complexity parameter to the full data (training + testing)

```{r}
heart_tree<-rpart()
...
```

- Find the Out of Bag (OOB) error for tree

```{r}

```


---

## Question 2: Bagging, Random Forest

- Compare the performance of classification trees (above), bagging, random forests for predicting heart disease based on the ``heart`` data.

- Use the training and testing sets from above. Train each of the models on the training data and extract the cross-validation (or out-of-bag error for bagging and Random forest). 


- For bagging use ``randomForest`` with ``mtry`` equal to the number of features (all other parameters at their default values). Generate the variable importance plot using ``varImpPlot`` and extract variable importance from the ``randomForest`` fitted object using the ``importance`` function.

```{r}
heart_bag<-randomForest()
# oob error rate
sum(heart_bag$err.rate[, 1])

varImpPlot()
importance()
```

- For random forests use ``randomForest`` with the default parameters. Generate the variable importance plot using ``varImpPlot`` and extract variable importance from the ``randomForest`` fitted object using the ``importance`` function.

```{r}
heart_rf <- randomForest()
#oob error rate
sum(heart_rf$err.rate[, 1])

varImpPlot()
importance()
```
---

# Question 3: Boosting

- For boosting use `gbm` with ``cv.folds=5`` to perform 5-fold cross-validation, and set ``class.stratify.cv`` to ``AHD`` (heart disease outcome) so that cross-validation is performed stratifying by ``AHD``.  Plot the cross-validation error as a function of the boosting iteration/trees (the `$cv.error` component of the object returned by ``gbm``) and determine whether additional boosting iterations are warranted. If so, run additional iterations with  ``gbm.more`` (use the R help to check its syntax). Choose the optimal number of iterations. Use the ``summary.gbm`` function to generate the variable importance plot and extract variable importance/influence (``summary.gbm`` does both). Generate 1D and 2D marginal plots with ``gbm.plot`` to assess the effect of the top three variables and their 2-way interactions. 

```{r}
heart_boost = gbm()

```


---


## Question 4: Gradient Boosting

Evaluate the effect of critical boosting parameters (number of boosting iterations, shrinkage/learning rate, and tree depth/interaction).  In ``gbm`` the number of iterations is controlled by ``n.trees`` (default is 100), the shrinkage/learning rate is controlled by ``shrinkage`` (default is 0.001), and interaction depth by ``interaction.depth`` (default is 1).

Note, boosting can overfit if the number of trees is too large. The shrinkage parameter controls the rate at which the boosting learns. Very small $\lambda$ can require using a very large number of trees to achieve good performance. Finally, interaction depth controls the interaction order of the boosted model. A value of 1 implies an additive model, a value of 2 implies a model with up to 2-way interactions, etc. the default is 1.


- Set the seed and train a boosting classification with ``gbm`` using 10-fold cross-validation (``cv.folds=10``) on the training data with ``n.trees = 5000``, ``shrinkage = 0.001``, and ``interaction.depth =1``. Plot the cross-validation errors as a function of the boosting iteration and calculate the test MSE.

```{r}
set.seed(301)
heart_boost = gbm(AHD ~ ., data = heart[train, ], distribution = "bernoulli", n.trees = 5000, interaction.depth = 1, shrinkage = 0.001, cv.folds = 10, class.stratify.cv = TRUE)

summary(heart_boost)

yhat_boost<-predict() #predict on test
#calculate MSE


```

- Repeat the above using the same seed and ``n.trees=5000`` with the following 3 additional combination of parameters: a) ``shrinkage = 0.001``, ``interaction.depth = 2``; b) ``shrinkage = 0.01``, ``interaction.depth = 1``; c) ``shrinkage = 0.01``, ``interaction.depth = 2``.

```{r}
set.seed(301)
heart_boost = gbm()


```


## Question 5: Extreme Gradient Boosting

Training an xgboost model with `xgboost` and perform a grid search for tuning the number of trees and the maxium depth of the tree. Also perform 10-fold cross-validation and determine the variable importance. Finally, compute the test MSE.

```{r}

train_control = trainControl(method = "cv", number = 10, search ="grid")

tune_grid<-  expand.grid(max_depth = c(1, 3, 5, 7), #larger makes model more complex and potentially overfit
                        nrounds = (1:10)*50, # number of trees
                        # default values below
                        eta = c(0.01,0.1,0.3), # learning rate (equivalent to shrinkage)
                        gamma = 0, #Minimum loss reduction required to make a further partition on a leaf node of the tree. The larger gamma is, the more conservative the algorithm will be.
                        subsample = 1,
                        min_child_weight = 1,
                        colsample_bytree = 0.6 #specify the fraction of columns to be subsampled
                        )

heart_xgb<-caret::train()

varimp <- varImp()
plot(varimp)

```

- Compare the the performance of the different models and summarize

# Deliverables

1. Questions 1-5 answered, pdf or html output uploaded to quercus
